### 2023-10-09 Analyzing perimetric data with the visualFields package/Analyzing perimetric data with the visualFields package.Rmd
Started with template.

---
documentclass: jss
author:
  - name: Cord Huchzermeyer
    orcid: 0000-0000-0000-0000
    affiliation: 'University Hospital Erlangen \AND Friedrich-Alexander-University Erlangen Nürnberg'
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{huchzi@email.de}
    url: https://huchzi.github.io
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Analyzing Perimetric data with the \\pkg{visualFields} package"
  # If you use tex in the formatted title, also supply version without
  plain:     "Analyzing Perimetric data with the visualFields package"
  # For running headers, if needed
  short:     "\\pkg{visualFields}: Statistics for Perimetry"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{R}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

# Introduction

# Psychophysical Thresholds

## Psychophysical Function

## Sensitivities, Defects, Probabilities

## Algorithms for Measuring Psychophysical Thresholds

# Mapping the Hill of Vision

## What is the Hill of Vision

## Global parameters

## Bebie-Curve

## Different patterns

## Plotting Visual Fields

# Technical details

## Humphrey

## Octopus

## Others

# Structure of perimetric data

# Importing perimetric data

# Setting up the Environment

# Statistical Analysis

# Plotting

# Conclusion

## Code formatting

In general, don't use Markdown, but use the more precise LaTeX commands instead:

* \proglang{Java}
* \pkg{plyr}

One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax).

If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute. You can also provide a custom identifier if necessary. See the header of Section \ref{r-code} for example.

# \proglang{R} code {short-title="R code" #r-code}

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```

## Features specific to \pkg{rticles} {short-title="Features specific to rticles"}

* Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409).
* Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format.